---
title: "Homework Template"
output: html_document
  html_document: default
  pdf_document: default
date: "2023-10-09"
---

```{r setup}
install.packages("ggplot2")
install.packages("palmerpenguins")
install.packages("janitor")
install.packages("dplyr")

library("ggplot2")
library("palmerpenguins")
library("janitor")
library("dplyr")
library("tinytex")

head(penguins_raw)
```

*The following is a template .rmd RMarkdown file for you to use for your homework submission.*

*Please Knit your .rmd to a PDF format or HTML and submit that with no identifiers like your name.*

*To create a PDF, first install tinytex and load the package. Then press the Knit arrow and select "Knit to PDF".*

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE}
penguins2 <- penguins

penguingraph <- ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm)) + geom_line()

penguingraph2 <- ggplot(penguins2, aes(x = body_mass_g)) +
  geom_line(aes(y = bill_length_mm, color = "Bill length"), linetype = "solid") +
  geom_line(aes(y = bill_depth_mm, color = "Bill depth"), linetype = "dashed") +
  labs(title = "How body mass affects measurement in mm of bill length and bill depth",
       x = "Mass of penguin",
       y = "Measurement in mm") +
  scale_color_manual(values = c("Bill length" = "darkgreen", "Bill depth" = "red"))

print(penguingraph2)
# There is no need to provide the code for your bad figure, just use echo=FALSE so the code is hidden. Make sure your figure is visible after you knit it. 

```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).
This graph is misleading as it includes a lot of data is on the graph as it has two y variables being used on the same y axis. This would confuse or mislead a viewer as it is a lot of information to pay attention to.
Also as there is two y variables and they are not close in values, it means that the values for bill depth are hard to see as they are made very small to fit in the values for bill length.
The points all joined by a line which does not give an indication of any correlation - whether this is positive or negative. Also by not displaying the individual data points you cannot see the spread of the data.
The colours are inappropriate as might not be accessible for any people with colour blindness.Also the dashed line makes it harder to see the values for bill depth. 
Bill length on the bottom of the key but at the top of the figure this would make this more confusing as you would need to pay attention to the colours and the key is in an illogical order.

*Include references.*

------------------------------------------------------------------------

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

```{r Data Exploration}
#install.packages(c("ggplot2", "palmerpenguins", "janitor", "dplyr"))

library(ggplot2)
library(palmerpenguins)
library(janitor)
library(dplyr)
library(ragg)

head(penguins_raw)
#loading data

write.csv(penguins_raw, "data/penguins_raw.csv")

source("functions/cleaning.r")
penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()
penguins_clean <- clean_function(penguins_raw)
penguins_clean <- as.data.frame(penguins_clean)

penguinexp <- ggplot(penguins_clean, aes(x = body_mass_g, y = culmen_length_mm, color = species)) +
  geom_point()
penguinexp

source("functions/plotting.r")

penguinexplor <- plot_exploratory_figure(penguins_clean)
penguinexplor

save_flipper_plot_png(penguins_clean, 
                      "figures/fig01_penguinexp.png", 
                      size = 15, res = 600, scaling=1)


# Make sure your code prints. 
```

First I have installed the packages needed to complete this analysis, these include ggplot (to form the graphs), palmerpenguins (this includes the data), janitor (provides functions that are needed to clean the data) and dplyr (gives functions that allow you to alter components of the data frame).

Then I checked whether the data frame I needed was properly installed by running the head() function which allows you to see the first few rows of the data frame and the column names.

Then I converted the penguin_raw data into a csv file.

For this analysis I have cleaned the code, meaning I have removed irrelevant columns such as comments from the data frame. 
The frame containing the function source() is to access the file that the functions I created are in, these functions include shorten_species. Doing this allows you to remove data that would impede the ability to make the graph and makes the data frame more readable.

I have then made exploratory graph depicting body mass (g) relationship to culmen length (mm) and colouring the different points by their species.
This was then saved to a file called "figures".

### Hypothesis

Alternative hypothesis or H1 is that penguin body mass increases as culmen length increases.And that this is different between species.
The null hypothesis or H0 is that there is no relationship between penguin body mass and penguin culmen length, and that there is no difference between species.


### Statistical Methods

Here I create a model exploring the relationship between body mass, culmen length and species.
I then check the assumptions of an ANCOVA analysis by doing a qqplot and plotting a residuals vs fitted plot.
The qq plot tests for a normal distribution, which the data does have.
The fitted vs residual plot tests for linearity.

Summary gives us estimates for the intercepts and slopes. As well, it gives us the p-values for these relationships so you can see if they're significant.
The function anova() completes an anova analysis on the model.

```{r Statistics}
pen_model <- lm(body_mass_g ~ species * culmen_length_mm, penguins_clean)

plot(pen_model, which = 2) 
plot(pen_model, which = 1) 

summary(pen_model)

anova(pen_model)

# Make sure your code prints. 

```
The adjusted R-squared value obtained for this is 0.785, which suggests that 78.5% of the variation is explained by this model. 
Body masses relationship to species and culmen length are both significant as the p-value obtained was <2e-16. The p value is 0.01549 for the interaction between body mass, species and culmen length, this value is below 0.05 so the interaction is significant.
The relationship between body mass and culmen length does differ between species.


### Results & Discussion

```{r Plotting Results}
penplot <- ggplot(penguins_clean, aes(x = body_mass_g , y = culmen_length_mm, colour = species)) +
  geom_point() +
  geom_smooth(method = "lm") +
    labs(x = 'Body mass (g)',  
      y = 'Culmen length (mm)',
      title = "Scatter Plot of Body Mass vs. Bill Length") +
  theme_bw()
penplot

penguinmodel <- plot_model_figure(penguins_clean)
penguinmodel

save_model_plot_png(penguins_clean, 
                      "figures/fig01_penguinmodel.png", 
                      size = 15, res = 600, scaling=1)

# Make sure your code prints. 
```

### Conclusion

A significant difference was found in the relationship between body mass and culmen length of different species when conducting the ANOVA test on the model as the p-value was equal to < 2e-16. There is also a statistically significant interaction between body mass and species as the p-value was also < 2e-16. This implies that the different penguin species are different sizes as they must have a different average body mass across species. The final ANOVA value implies that there is a significant interaction effect (p = 0.0154) between body mass and species culmen length which suggests that the relationship between culmen length and body mass varies among different species. 
The values for the summary() function also tell us about the relationship between body mass and culmen length in the different species. For example when comparing Adelie and Chinstrap (speciesChinstrap:culmen_length_mm) the relationship of the Chinstrap penguins is weaker as the value is -35.38, it is also statistically significant as the p-value is 0.047. The difference in the relationship of body mass and culmen length is positive, which would suggest a stronger relationship compared to Adelie. However, the p-value tells us that this is not statisitically significant as the p-value is 0.344.

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*

